---
title: "02_11_19"
author: "Aakash Ahuja"
date: "11/02/2020"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
always_allow_html: yes
---
**Loading Libraries and Data**
```{r}
#Loading Libraries
#install.packages('tidycensus')
library(tidycensus)
census_api_key("38db8e5473d1110e8747ea3c3607201b634f3538")
#install.packages('readxl')
library(readxl)
#install.packages('plotly')
library(plotly)
#install.packages('lubridate')
library(lubridate)
library(tidyr)
library(dplyr)
#install.packages('forecast')
library(forecast)
library(scales)
#Loading the State abbreviations Data
state_abbreviations<-read.table('state_abbrevs.txt', header = T)

#Loading the US State Data
state_and_us_data<-read_xls('State_and_US_SA.xls')

#Slicing the extra rows and columns from the data
t1<-state_and_us_data %>% slice(5)
state_and_us_data<-state_and_us_data %>% slice(6:545)
colnames(state_and_us_data)<-t1

#Loading the CPI Data
cpi<-read.csv('cpi.csv')
cpi<-subset(cpi,Year>=1975 & Year<=2019)
```

**Adjusting the Month Variable**
```{r}
str(state_and_us_data$Month)
#Converting Date Variable
temporary_date<-strsplit(as.character(state_and_us_data$Month), "M")
temporary_date<- as.data.frame(matrix(unlist(temporary_date), byrow = T, ncol = 2))

temporary_date$V1<- as.numeric(as.character(temporary_date$V1))
temporary_date$V2<- as.numeric(as.character(temporary_date$V2))

temporary_date$Date <- with(temporary_date, sprintf("%d-%02d", V1, V2))

temporary_date$Date<-parse_date_time(temporary_date$Date,"ym")

#Putting this transformed date variable into main date variable
state_and_us_data$Month<-temporary_date$Date
str(state_and_us_data$Month)
```

```{r}
#Adjusting House Price Index for inflation after using the CPI values
state_and_us_data$cpi<-cpi$Value

state_and_us_data$inflation_adjusted_us_data<-(as.numeric(state_and_us_data$`United States seasonally adjusted`) * 174.600)/state_and_us_data$cpi
```


```{r}
#Converting to time series data
myts<-ts(state_and_us_data$`United States seasonally adjusted`,start = c(1975,1), end = c(2019,12), frequency = 12)

#Converting to time series non seasonally adjusted data
myts1<-ts(state_and_us_data$inflation_adjusted_us_data,start = c(1975,1), end = c(2019,12), frequency = 12)

#Plotting my time series
plot(myts, col='red', xlab="Year", ylab="House Price Index")+
lines(myts1, col='Blue')
legend('topright',legend = c("HPI", "HPI(Adjusted)"), col=c("red",'Blue'))
```
After adjusting for inflation, it can be seen that the house prices increased over the span of 45 years(1975-2019). Although, there is a sudden drop in house prices observed during the time period 2008-2010. Then the house prices start increasing again after year 2012


```{r}
#Using Gather function to prepare longer version of us state data
c1<-colnames(state_and_us_data)[2:52]
temp2<-state_and_us_data[,-c(53,54)]
state_and_us_data_long<-temp2 %>%
  gather(key="Code",value = "hpi_value",c1 )

state_and_us_data_long<-left_join(state_and_us_data_long,state_abbreviations,by='Code')
state_and_us_data_long<-state_and_us_data_long[,-c(2)]
```

To observe the variations in house price index on the basis of the states, we create a new data frame by aggregating the longer version of us data and the mean of hpi values for every state and year.

```{r}
#By state HPI
str(state_and_us_data_long$hpi_value)
Mean_House_Price_By_State<-aggregate(as.numeric(hpi_value)~year(Month)+State,FUN =mean, data = state_and_us_data_long)
colnames(Mean_House_Price_By_State)<-c("Month","State","hpi_value")

#Plotting HPI for each state
temp1<-Mean_House_Price_By_State
temp1$Month<-as.Date(as.character(Mean_House_Price_By_State$Month), format = "%Y")

ggplot(data = temp1, aes(x=Month,y=hpi_value))+
  geom_point(alpha=0.5, size=0.2 )+
  geom_smooth(method = 'loess')+
  facet_wrap(~State,ncol=13)+ scale_x_date(labels = date_format("%y"))


#Calculating  Percentage change in House Price Index for each state by year
Mean_House_Price_By_State$percentage_change<-0

for (i in 1:nrow(Mean_House_Price_By_State)) {
  if (Mean_House_Price_By_State$Month[i]!=1975){
  Mean_House_Price_By_State$percentage_change[i]<-((Mean_House_Price_By_State$hpi_value[i]-Mean_House_Price_By_State$hpi_value[i-1])/(Mean_House_Price_By_State$hpi_value[i]))*100
  }
  else{
    Mean_House_Price_By_State$percentage_change[i]<-0
  }
}

Mean_House_Price_By_State[which.max(Mean_House_Price_By_State$percentage_change),]
#Rhode Island in year 1987

Mean_House_Price_By_State[which.min(Mean_House_Price_By_State$percentage_change),]
#Nevada in the year 2009
```
It can be seen that every state in US follows a general trend, in which the house price index tends to increase over time but there's a sudden drop in the house prices in 2008 or 2009 due to the economic recession. The biggest change of 24.65% in house prices was observed for Rhode Island for the year 1987 and the smallest change of -38% was observed for Nevada for the year 2009. Although there are states like Nevada and District of Columbia, which deviates from the similar pattern as the house prices keeps on increasing even after the recession period of 2008-2010. These two states can be considered as outliers.


```{r}
#By Region HPI
Mean_House_Price_By_Region<-aggregate(as.numeric(hpi_value)~year(Month)+Region,FUN = mean, data = state_and_us_data_long)
colnames(Mean_House_Price_By_Region)<-c("Month","Region","House_Price_Index")
#Plotting HPI for each region
temp2<-Mean_House_Price_By_Region
temp2$Month<-as.Date(as.character(Mean_House_Price_By_Region$Month), format = "%Y")

ggplot(data = temp2, aes(x=Month,y=House_Price_Index))+
  geom_point(alpha=0.5, size=0.2 )+
  geom_smooth(method = 'loess')+
  facet_wrap(~Region,ncol=2)+ scale_x_date(labels = date_format("%y"))

```
After plotting the region-wise distribution of house prices, it is seen that it follows the general trend and the house prices increase over the period, with a sudden drop in house prices during the period 2008-2010.


```{r}
present_density <- get_estimates(geography = "state", product = "population")
present_density <- present_density[53:103,]
```

```{r}

temp1<-aggregate(as.numeric(hpi_value)~State,FUN =mean, data = state_and_us_data_long)
colnames(temp1)<-c("State","House_Price_Index")

#Plotting HPI for each state

library(scales)

g<-ggplot(data = temp1, aes(x=State,y=House_Price_Index))+ geom_bar(stat = "identity")
ggplotly(g)
names(present_density)[1] <-"State"
temp3<-left_join(Mean_House_Price_By_State,present_density,by = "State")

temp4 <-aggregate(as.numeric(hpi_value)~State,FUN =mean, data = temp3)

temp4 <- cbind(temp4,present_density$value)


colnames(temp4)<-c("State","House_Price_Index", "Density")
fit1 <- lm(House_Price_Index~Density, data = temp4)
summary(fit1)
cor(temp4$House_Price_Index,temp4$Density) 
ggplot(data = temp4, aes(y=Density, x = House_Price_Index))+geom_point() + geom_smooth(method='lm',se=F) 

temp4['Density'>10000]
temp4 <- temp4[-c(9),]
print(temp4)
ggplot(data = temp4, aes(y=Density, x = House_Price_Index))+geom_point() + geom_smooth(method='loess',se=F) 
cor(temp4$House_Price_Index,temp4$Density)


t5 <- state_abbreviations
t5 <- t5[-c(9),]
t5$Density <- present_density$value[which(present_density$value<10000)]
t5$hpi <- temp4$House_Price_Index


ggplot(data = t5, aes(x=Density,y=hpi))+
  geom_point(alpha=0.5, size=0.2 )+
  geom_smooth(method = 'loess')+
  facet_wrap(~Region,ncol=2)

cor(t5[t5$Region=='Midwest',]$Density,t5[t5$Region=='Midwest',]$hpi)
cor(t5[t5$Region=='Northeast',]$Density,t5[t5$Region=='Northeast',]$hpi)
cor(t5[t5$Region=='South',]$Density,t5[t5$Region=='South',]$hpi)
cor(t5[t5$Region=='West',]$Density,t5[t5$Region=='West',]$hpi)

#NumericVariables<-apply(t5$Density,Outlier_capping)
```


```{r}

pop_1990 <- get_decennial(geography = "state", variables = "P0010001", year = 1990)
pop_2000 <- get_decennial(geography = "state", variables = "P001001", year = 2000)
pop_2010 <- get_decennial(geography = "state", variables = "P001001", year = 2010)
pop_2010 <- pop_2010[-c(52),]

mean_hpi_1990 <- subset(Mean_House_Price_By_State,Month >=1990 & Month<=2000)
mean_hpi_1990<-aggregate(hpi_value~State,FUN =mean, data = mean_hpi_1990)
print(mean_hpi_1990)

mean_hpi_2000 <- subset(Mean_House_Price_By_State,Month >=2000 & Month<=2010)
mean_hpi_2000<-aggregate(hpi_value~State,FUN =mean, data = mean_hpi_2000)
print(mean_hpi_2000)

mean_hpi_2010 <- subset(Mean_House_Price_By_State,Month >=2000 & Month<=2019)
mean_hpi_2010<-aggregate(hpi_value~State,FUN =mean, data = mean_hpi_2010)
print(mean_hpi_2010)

hpi_pop_1990 <- cbind(pop_1990,hpi_mean = mean_hpi_1990$hpi_value)
hpi_pop_2000 <- cbind(pop_2000,hpi_mean = mean_hpi_2000$hpi_value)
hpi_pop_2010 <- cbind(pop_2010,hpi_mean = mean_hpi_2010$hpi_value)
hpi_pop_1990$year_range <- "1990-2000"
hpi_pop_2000$year_range <- "2000-2010"
hpi_pop_2010$year_range <- "2010-Present"

hpi_pop_all_years <- rbind(hpi_pop_1990,hpi_pop_2000,hpi_pop_2010) 
ggplot(data = hpi_pop_all_years, aes(x=value,y=hpi_mean))+
  geom_point(alpha=0.5, size=0.2 )+
  geom_smooth(method = 'loess')+
  facet_wrap(~year_range,ncol=3)

```
It can be seen from the three graphs that during the 1990-2000 period, the house prices were extremely low even for higher populated states. Whereas, the trend is sort of similar in the next time span of 20 years(2000-2019), where the house prices increased even for states in which the population was extremely low.



